Annotating, understanding, and predicting long-term video memorability

### 2018-01-15 ARTICLE.tex
Added Modelling part of the paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MEMORABILITY PREDICTION} %for Khartik
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{MEMORABILITY PREDICTION} %for Khartik
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Until now, we have seen how we collected data and memorability annotations from users.
We then computed memorability scores and analyzed how different users fared in remembering the video sequences from the movies they have already watched.
In this section, we move towards building a machine learning model that can learn and then predict the memorability scores of a video from its audio-visual features. 
We pose the problem of predicting memorability scores as a standard regression problem and Figure \ref{prop-apprch} illustrates the different steps in our method.
From the figure, we can see that the video sequences are converted into feature vectors and then a regression model is trained that can in-turn predict the memorability scores for new video sequences.
In this following sub-sections, we explain our choice of features, models to address the problem in hand.

\begin{figure}[h]	  
  \centering
    \includegraphics[width=0.9\columnwidth]{figures/approach.pdf}
		\caption{Proposed approach for automatically predicting memorability scores.}
    \label{prop-apprch}
\end{figure}

Before we go into the technical details about the different audio-visual features and regression models we explored, we will talk about how we split our dataset.
We split our dataset, at the level of movies, into training (70\%), validation (15\%) and test (15\%) data, which translates into 70 movies in the training set and 15 movies each in the validation and test sets.
We chose to split our dataset at the level of movies, instead of the video sequences, in order to avoid sequences from the same movie being present in the train as well as the evaluation set (validation+test).

\subsection{Feature extraction}
\label{feat-extr}
We explore a variety of audio-visual features in building a model that can predict memorability scores for new video sequences.
We mostly stick to state-of-the-art deep learning features that have been proposed in vision and audio research communities for tasks like video classification \cite{c3d-feat}, event detection \cite{audioset-feat}, image captioning \cite{caption-feat}, and concept detection \cite{sb-feat}.

We use the features extracted from the C3D model, a 3-dimensional convolutional network proposed for generic video analysis \cite{c3d-feat}. 
The main motivation to use these features is that these features try to encode the motion information in the video.
The model has been proposed for video analysis and is not an extension of a model for image analysis, unlike other state-of-the-art models like VGG16 \cite{vgg16}.
We use the model trained on the Sports-1M dataset \cite{c3d-feat} that is provided by the authors of the paper.
We explore the output of the 3 fully connected layers of the network: fc6, fc7, and fc8 with a dimensionality of 4096.
Since the dimensionality of the features is very high, we also explore a dimension reduction method (Principal Component Analysis) for modelling.

Google Inc. recently released a large-scale dataset of manually annotated events: AudioSet \cite{audioset-feat}. 
The dataset consists of 10-second audio clips taken from YouTube videos along with code to extract 128-dimensional embeddings for your own audio clips.
We use these features as they are state-of-the-art in the audio event detection research and events could play a major role in how people remember sequences in movies.
We extract the 128-dimensional embeddings for our dataset and use these embeddings in training the regression models.

Now, we move onto features related to emotion.
Emotion and memory are correlated \cite{}, which is why we look at emotion related features in building a model for memorability prediction.
For emotion from visual content, we resort to a visual sentiment concept detector: Sentibank \cite{sb-feat}.
SentiBank is a set of 1200 trained visual concept detectors providing a mid-level representation of sentiment from visual content.
We use the binary code for concept detection, from images, provided by the authors.
We sample one frame for every second of the video sequence in our dataset, resulting in 10 frames per video sequence.
For each of these 10 frames, we run the sentibank concept detector and obtain two pieces of information: concepts with probabilities and features.
Concepts are adjective--noun pairs and the probability represents how likely this concept is depicted visually in that particular frame.
We rank the concepts based on the probability of their occurrence in the frame and take the top-50 concepts.
For each of the 50 concepts, we extract 300-dimensional word2vec \cite{word2vec} embeddings by training on english wikipedia corpus and taking an average of all the vectors to obtain a single vector per frame.
We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
Sentibank detectors also provide 4096-dimensional features for each frame and we take the average across all the frames to obtain one 4096-dimensional feature vector for each video.
In the end, we use two features: 300-dimensional concept vectors and 4096-dimensional feature vectors as features to build a model for memorability prediction.
Emotion will also be embedded in the audio signal and hence we resort to an audio-visual analysis of the video sequence to obtain arousal and valence scores \cite{affect}.
Arousal is the dimension of emotion that measures the excitement in the video, while valence measures whether the video happy or sad, following a circumplex model of affect \cite{affect-model}.
For each frame in the video sequence, we compute the arousal and valence scores using the methods proposed in \cite{affect}.
We concatenate the arousal and valence scores for the first 200 frames in each video sequence resulting in a 400-dimensional feature vector (200 for arousal and 200 for valence) for a video.

In addition to emotion, semantics also play a role in memory.
We utilize the ongoing research in image captioning to capture the semantics \cite{caption-feat}.
We use the code provided by the authors to extract captions for videos in our dataset.
We sample one frame for every second of the video sequence in our dataset, resulting in 10 frames per video sequence.
For each of these 10 frames, we run the caption detector and obtain a caption for the frame.
For each word in the caption, we extract 300-dimensional word2vec \cite{word2vec} embeddings by training on english wikipedia corpus and taking an average of all the words to obtain a single vector per frame.
We repeat this process for all the 10 frames and take the average of all the vectors to obtain a single feature vector for each video.
We use these 300-dimensional vectors for building our model.

\subsection{Modelling and evaluation}
\label{model-eval}
We discussed the variety of features that we extract from the audio-visual content.
Now, we turn to picking a regression model that can effectively predict memorability scores for videos in our dataset.
We choose to go with the Support vector regression model (SVR) \cite{svr}, with a linear and an RBF kernel, for its popularity and wide applicability.
Since we are posing the memorability prediction problem as a regression, we use the standard regression metric: Mean squared error ($mse$) for evaluation.
Additionally, we also use the spearmann correlation ($spcorr$) to measure the rank correlation between the predicted memorability scores and the ground-truth.
This metric gives us an indication of the correlation between the predicted memorability scores and the human annotated memorability scores.

\section{Memorability prediction results}
\label{pred-res}
In this section, we will discuss how the models trained on different features perform in predicting memorability scores of new video sequences.
Each of the video sequences in our dataset has multiple annotations and we pick sequences with atleast 4 annotations for training the models.
In our dataset, 660 sequences (need to confirm) out of the 700 sequences have atleast 4 annotations.
One set of results correspond to using these 660 sequences. 
Another set of results correspond to using 330 sequences (need to confirm) which have atleast 10 annotations.

Since we are using an SVR with linear and rbf kernels, we train the models on the training set (70 movies) and validate the model on the validation set (15 movies).
We pick the model parameters: kernel (linear or rbf), $C$ and $\gamma$ that give the best performance in terms of $mse$ and $spcorr$.
We retrain the model with these parameters and evaluate on the final test set.
In the experiments where we use a dimension reduction method (PCA): C3D and SentiBank features, we try to retain 95\% variance in the data while reducing the feature dimensions.
Tables \ref{res-4-ann} and \ref{res-10-ann} report the memorability prediction results for different features using models trained on sequences with atleast 4 and 10 annotations.
Observing the tables, you can clearly see that the image captions perform the best among all the features.
These features capture the semantics in the video and are helping in a better prediction of the memorability scores.
The next best performance (apart from the combination of all the features) is observed for the features from the AudioSet embeddings.
Since these embeddings encode the event related information, we can say that the events happening in a video play a significant role in how the video is remembered.
A combination of all the features is the second best performing feature combination in terms of the metrics.
We concatenate all the features into a single feature vector and train a model which is then used to predict the memorability scores on the test data. 
We observe that there is a decline in performance when compared to the best performing features: image captions and this is because the other features are hindering the performance of the regression model.

One of our initial hypotheses was that emotion would play a big role in memorability of a video sequence, supported by literature from psychology.
We used different set of features to encode the emotion in a video: Affect \cite{affect} and SentiBank \cite{sb-feat}.
Observing the tables, you can say that the SentiBank features perform reasonably well, but it is not close to the performance of features encoding the semantics.
This could be because of the choice of features that encode emotion or the performance of the emotion models not being good enough for such an application.
From the analysis on this dataset and the features we used for encoding emotion, we cannot conclude that emotion plays an important role in memorability.

Comparing the results across the two tables, we observe that the model trained on sequences with atleast 4 annotations performs better than the model trained on sequences with atleast 10 annotations.
The only exception being the C3D feature, where the model trained on sequences with 10 annotations performs marginally better than the model trained on 4 annotations. 
Now, we will investigate how the performance changes with increasing number of annotations: from 4 to 15.
For this, we take the best performing feature: image captions and analyze its performance with models trained on sequences with different number of annotations.
We take sequences with 4 annotations and keep increasing the number of annotations per sequence to 15.
In each step, we repeat the training-validation-testing procedure of the previous experiment and compute the evaluation metrics.
We observed that the $mse$ does not change significantly in each of the steps and hence we provide an demonstration of how $spcorr$ varies with increasing number of annotations.
Figure \ref{num-ann} demonstrates the result of this experiment, where we observe that beyond 8 annotations the value of $spcorr$ keeps going down.

\begin{table*}[!ht]
\centering
    \begin{tabular}{| c | c | c | c |}
    \hline
				Feature& Model& $mse$& $spcorr$\\ \hline
				C3D (4096)& Linear, $C$=1.0& 0.08& 0.26\\ \hline
				C3D (PCA) (225)& Linear, $C$=1.0& 0.08& 0.21\\ \hline
				AudioSet (128)& RBF, $C$=1.0, $\gamma$=0.01& 0.06& 0.22\\ \hline
				Affect (400)& RBF, $C$=1.0, $\gamma$=0.1& 0.07& 0.17\\ \hline
				SentiBanl Conc. (300)& RBF, $C$=1.0, $\gamma$=10& 0.08& 0.13\\ \hline
				SentiBank Feat. (4096)& RBF, $C$=1.0, $\gamma$=0.1& 0.07& 0.21\\ \hline
				SentiBank Feat. (PCA) (225)& RBF, $C$=1.0, $\gamma$=0.01& 0.07& 0.21\\ \hline
				\textbf{Captions (300)}& \textbf{RBF, $C$=1.0, $\gamma$=0.01}& \textbf{0.06}& \textbf{0.31}\\ \hline
				All (1578)& RBF, $C$=1.0, $\gamma$=0.1& 0.06& 0.23\\ \hline
    \end{tabular}		
		\caption{Regression results for different features with models trained on sequences that have atleast 4 annotations.}
		\label{res-4-ann}	
\end{table*}

\begin{table*}[!ht]
\centering
    \begin{tabular}{| c | c | c | c |}
    \hline
				Feature& Model& MSE& SpearCorr\\ \hline
				C3D (4096)& Linear, $C$=1.0& 0.08& 0.24\\ \hline
				C3D (PCA) (225)& Linear, $C$=1.0& 0.09& 0.23\\ \hline
				AudioSet (128)& RBF, $C$=1.0, $\gamma$=0.1& 0.1& 0.15\\ \hline
				Affect (400)& RBF, $C$=1.0, $\gamma$=0.01& 0.09& 0.17\\ \hline
				SentiBanl Conc. (300)& Linear, $C$=1.0& 0.08& 0.14\\ \hline
				SentiBank Feat. (4096)& RBF, $C$=1.0, $\gamma$=10& 0.09& 0.14\\ \hline
				SentiBank Feat. (PCA) (225)& Linear, $C$=1.0& 0.08& 0.03\\ \hline
				\textbf{Captions (300)}& \textbf{RBF, $C$=1.0, $\gamma$=0.01}& \textbf{0.06}& \textbf{0.22}\\ \hline
				All (1578)& RBF, $C$=, $\gamma$=& 0.07& 0.21\\ \hline
    \end{tabular}		
		\caption{Regression results for different features with models trained on sequences that have atleast 10 annotations.}
		\label{res-10-ann}	
\end{table*}

\begin{figure}[h]	  
  \centering
    \includegraphics[width=0.7\columnwidth]{figures/annotations.pdf}
		\caption{Spearmann correlation on test data with models trained using features from image captions for varying number of annotations.}
    \label{num-ann}
\end{figure}

\subsection{Cross-validation}
\label{cross-val}
Since we have a relatively small dataset, we perform a cross-validation experiment in order to ensure that the model is not overfitting.
We perform a 10-fold cross-validation on the cross-validation set: combining the training and validation sets.
We report the results (both $mse$ and $spcorr$) of the cross-validation for the following parameters: features from image captions, model trained on sequences with atleast 4 annotations, RBF kernel.
Figure \ref{cross-val} reports the results of the cross-validation experiment and we observe that the evaluation metrics do not change much across the different folds.
We can conclude that the trained model is generalizable and there is not much overfitting.

\begin{figure}[h]	  
  \centering
    \includegraphics[width=0.8\columnwidth]{figures/cross-val.pdf}
		\caption{Cross-validation performance on a model trained on sequences with atleast 4 annotations with features from image captions and an RBF kernel.}
    \label{cross-val}
\end{figure}
